---
title: "Clustering Spotify songs with K-Means"
author: "Rahma Fairuz Rania"
date: '2022-07-09'
output: 
 html_document:
   toc: true
   toc_float: true
   highlight: zenburn
   df_print: paged
   theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Case : From music data in Spotify, we want to clustering music with the same charateristic in 1 group. Data can be downloaded here https://www.kaggle.com/zaheenhamidani/ultimate-spotify-tracks-db

# Read Data

```{r}
songs <- read.csv('SpotifyFeatures.csv')
head(songs)
```

# Data Wrangling

Check data structure

```{r}
str(songs)
```

From data songs, there's column below

* `ï..genre` : Genre of Music
* `artist_name` : The Name of Music's Singer
* `track_name` : Name of the Song
* `track_id` : The Spotify ID for the track.
* `popularity` : Song Popularity
* `acousticness` : A confidence measure from 0.0 to 1.0 of whether the track is acoustic. 
* `danceability` : Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity. 
* `` **duration_ms** : The duration of the track in milliseconds.
* `energy` : Energy is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity.
* `instrumentalness` : Predicts whether a track contains no vocals. "Ooh" and "aah" sounds are treated as instrumental in this context.
* `key` : The key the track is in. Integers map to pitches using standard Pitch Class notation. E.g. 0 = C, 1 = C♯/D♭, 2 = D, and so on. If no key was detected, the value is -1.
* `liveness` : Detects the presence of an audience in the recording.
* `loudness` : The overall loudness of a track in decibels (dB). Loudness values are averaged across the entire track and are useful for comparing relative loudness of tracks. 
* `mode` :  Mode indicates the modality (major or minor) of a track, the type of scale from which its melodic content is derived.
* `speechiness` : Speechiness detects the presence of spoken words in a track.
* `tempo` : The overall estimated tempo of a track in beats per minute (BPM).
* `time_signature` : The time signature (meter) is a notational convention to specify how many beats are in each bar (or measure).
* `valence` : A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track.


```{r, warning=FALSE, message=FALSE}
library(dplyr)
songs <- songs %>% mutate(duration_ms = duration_ms/60000) %>% rename(duration_in_min = duration_ms)

head(songs)
```

Check Missing Value If exist, remove it

```{r}
anyNA(songs)
```

Check duplicate data. If exist, remove it

```{r}
sum(duplicated(songs$track_id))
spotify <- songs[!duplicated(songs$track_id),]

sum(duplicated(spotify$track_id))
```

# Exploratory Data Analysis

We want to do clustering from the track_id Assign track_id column into rownames

```{r}
rownames(spotify) <- spotify$track_id 
```

Select song with popularity >= 70 to speed up computation

```{r}
spotify <- spotify %>% filter(popularity >= 70)
```

Because K-Means use the numerical variable, do subset data just for numeric

```{r}
spotify <- spotify %>% select_if(is.numeric)
```

Check the numerical variable

```{r}
str(spotify)
```

# Data Pre-Processing

Do scaling data for standarization

```{r}
spotify_scale <- scale(spotify)
```


Search for optimum k with fviz_nbclust method from library(factoextra)

```{r, warning=FALSE, message=FALSE}
RNGkind(sample.kind = "Rounding")
set.seed(572)
library(factoextra)

fviz_nbclust(x = spotify_scale,
             FUNcluster = kmeans,
             method = "wss")
```

From the graph above, let's take 4 for the optimum k to do k-means clustering

# K-Means Clustering

Do the clustering using the optimum key which we have search it though elbow method above.

```{r, warning=FALSE, message=FALSE}
RNGkind(sample.kind = "Rounding")
set.seed(100)
spotify_kmeans <- kmeans(spotify_scale, centers = 4)

spotify_kmeans$size
```
From data above, we know that in cluster 1 there's 690 observation, cluster 2 there's 1396 observation, cluster 3 there's 1065 observation, cluster 4 there's 684 observation. For better information, let's do profiling with **ggRadar** function from library **ggiraphExtra**

```{r, warning=FALSE, message=FALSE}
# Assign clustering result into data
spotify$cluster <- spotify_kmeans$cluster

library(ggiraphExtra)
ggRadar(data = spotify, mapping = aes(colours = cluster), interactive = T)
```

For better information, we can use summarize for viewing the minimum and maximum value of each cluster

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
spotify_centroid <- spotify %>% group_by(cluster) %>% summarise_all(mean) 

spotify_centroid %>% 
  pivot_longer(-cluster) %>% 
  group_by(name) %>% 
  summarize(
    min = which.min(value),
    max = which.max(value))
```



# Conclusion

From information above, here are some insight that can help to make decision

* `Cluster 1`  : 690 observation
* `Cluster 2`  : 1396 observation
* `Cluster 3`  : 1065 observation
* `Cluster 4`  : 684 observation.

Audio characteristics in every cluster

* `Cluster 1` 
  Highest : acousticness, instrumentalness; 
  Lowest  : energy, liveness, loudness, speechiness, tempo, valence
* `Cluster 2` 
  Highest : loudness, valence; 
  Lowest  : not spesific
* `Cluster 3` 
  Highest : duration_in_min, energy, liveness, tempo; 
  Lowest  : acousticness, danceability, popularity
* `Cluster 4` 
  Highest : danceability, popularity, speechiness; 
  Lowest  : duration_in_min, instrumentalness